---
title: "results04_human-llm_accuracy-stats"
author: "Tommy Botch"
date: "2024-11-15"
output: html_document
---

# Setup environment

## Load packages and setup paths for analysis

Set up directories - since we're doing this locally (and mounting the server), 
we're going to use a different base_dir

```{r setup, include=FALSE}

# Source the utils.R file to load the functions
source("../utils/utils.R")

# Load required and additional packages
install_and_load()

# Set the paths for this notebook
set_paths(analysis_name="behavioral", plots_name="human-llm-comparison")
```

# Human-LLM accuracy analyses

## Data loading

Load data and set factors. Our columns are the following:
- **modality:** which modality was the stimulus presented as (video, audio, text, or model name) 
- **word_index:** the word index within the story
- **task:** which of the three stories was the data from (black, wheretheressmoke, howtodraw)

We create a contrast of modality such that we test the hypothesis that multimodal (video) > spoken (audio) > written (text).

```{r}

results_fn <- file.path(results_dir, "all-task_group-analyzed-behavior_window-size-00025_human-model-lemmatized.csv")

# Load data and convert columns to factors
df_results <- load_and_prepare_data(
  file_path=results_fn,
  factor_columns=c('modality', 'word_index', 'task')
)

# ADD A COLUMN THAT GROUPS HUMANS/MODELS
df_results <- df_results %>%
  mutate(model_type = case_when(
    modality %in% c("video") ~ "video",
    modality %in% c("audio") ~ "audio",
    modality %in% c("text") ~ "text",
    modality %in% c('roberta', 'electra', 'xlm-prophetnet') ~ "mlm",
    TRUE ~ "clm"  # Default case
  ))

# Set the levels for the modality column
df_results <- set_factor_levels(
  df=df_results,
  column_name="model_type",
  levels=c("video", "audio", "text", "clm", "mlm")
)

df_results <- create_column_contrast(
  df=df_results, 
  column_name="model_type",
  n_poly=2
)
```

## Analysis 1A - Binary accuracy

We first investigate whether humans exhibit an advantage of binary accuracy (0/1) over LLMs based on the modality of the stimulus. We use a logistic mixed-effects model to classify accuracy based on modality.

```{r}

binary_accuracy_formula <- 'accuracy ~ model_type + (1|modality) + (1|word_index:task)'

# Fit the logistic mixed-effects model
model <- glmer(
  binary_accuracy_formula, 
  data=df_results, 
  family=binomial(link="logit")
)

tab_model(model, show.se=TRUE, show.stat=TRUE, show.r2=TRUE, show.df=TRUE)

# Display results of fixed effects
display_model_summary(model)

# Conduct pairwise comparisons between modalities
pairwise_comparisons <- emmeans(model, pairwise ~ model_type)
display_pairwise_comparisons(pairwise_comparisons)
```
## Analysis 1B - Task-level comparisons

```{r}

binary_accuracy_formula <- 'accuracy ~ model_type + (1|modality) + (1|word_index)'

```

### Story = Black

```{r}

model <- lmer_task(
  task_name='black', 
  df=df_results, 
  formula=binary_accuracy_formula,
  family=binomial(link="logit")
)

# Output the results
tab_model(model, show.se = TRUE, show.stat = TRUE, show.r2 = TRUE, show.df=TRUE)

# Display results of fixed effects
display_model_summary(model)

# Conduct pairwise comparisons between modalities
pairwise_comparisons <- emmeans(model, pairwise ~ model_type)
display_pairwise_comparisons(pairwise_comparisons)
```

### Story = Smoke

```{r}

model <- lmer_task(
  task_name='wheretheressmoke', 
  df=df_results, 
  formula=binary_accuracy_formula,
  family=binomial(link="logit")
)

# Output the results
tab_model(model, show.se = TRUE, show.stat = TRUE, show.r2 = TRUE, show.df=TRUE)

# Display results of fixed effects
display_model_summary(model)

# Conduct pairwise comparisons between modalities
pairwise_comparisons <- emmeans(model, pairwise ~ model_type)
display_pairwise_comparisons(pairwise_comparisons)
```

### Story = Draw

```{r}

model <- lmer_task(
  task_name='howtodraw', 
  df=df_results, 
  formula=binary_accuracy_formula,
  family=binomial(link="logit")
)

# Output the results
tab_model(model, show.se = TRUE, show.stat = TRUE, show.r2 = TRUE, show.df=TRUE)

# Display results of fixed effects
display_model_summary(model)

# Conduct pairwise comparisons between modalities
pairwise_comparisons <- emmeans(model, pairwise ~ model_type)
display_pairwise_comparisons(pairwise_comparisons)
```

## Analysis 2A - Continuous accuracy

```{r}

continuous_accuracy_formula <- 'fasttext_top_word_accuracy ~ model_type + (1|modality) + (1|word_index:task)'

# Fit the logistic mixed-effects model
model <- lmer(
  continuous_accuracy_formula, 
  data=df_results, 
)

tab_model(model, show.se=TRUE, show.stat=TRUE, show.r2=TRUE, show.df=TRUE)

# Display results of fixed effects
display_model_summary(model)

# Conduct pairwise comparisons between modalities
pairwise_comparisons <- emmeans(model, pairwise ~ model_type)
display_pairwise_comparisons(pairwise_comparisons)

```

## Analysis 2B - Task-level comparisons

```{r}

continuous_accuracy_formula <- 'fasttext_top_word_accuracy ~ model_type + (1|modality) + (1|word_index)'

```

### Story = Black

```{r}

model <- lmer_task(
  task_name='black', 
  df=df_results, 
  formula=continuous_accuracy_formula,
)

# Output the results
tab_model(model, show.se = TRUE, show.stat = TRUE, show.r2 = TRUE, show.df=TRUE)

# Display results of fixed effects
display_model_summary(model)

# Conduct pairwise comparisons between modalities
pairwise_comparisons <- emmeans(model, pairwise ~ model_type)
display_pairwise_comparisons(pairwise_comparisons)
```

### Story = Smoke

```{r}

model <- lmer_task(
  task_name='wheretheressmoke', 
  df=df_results, 
  formula=continuous_accuracy_formula,
)

# Output the results
tab_model(model, show.se = TRUE, show.stat = TRUE, show.r2 = TRUE, show.df=TRUE)

# Display results of fixed effects
display_model_summary(model)

# Conduct pairwise comparisons between modalities
pairwise_comparisons <- emmeans(model, pairwise ~ model_type)
display_pairwise_comparisons(pairwise_comparisons)
```

### Story = Draw

```{r}

model <- lmer_task(
  task_name='howtodraw', 
  df=df_results, 
  formula=continuous_accuracy_formula,
)

# Output the results
tab_model(model, show.se = TRUE, show.stat = TRUE, show.r2 = TRUE, show.df=TRUE)

# Display results of fixed effects
display_model_summary(model)

# Conduct pairwise comparisons between modalities
pairwise_comparisons <- emmeans(model, pairwise ~ model_type)
display_pairwise_comparisons(pairwise_comparisons)
```