---
title: "human_accuracy_stats"
author: "Tommy Botch"
date: "2024-11-15"
output: html_document
---

# Setup environment

## Load packages

```{r}
# remove all variables
rm(list=ls()) 

# install.packages(c("lmerTest", "emmeans", "glue", "ggeffects", "ggplot2", 
#                    "ggsignif", "ggpattern", "viridis", "forcats", "dplyr",
#                    "stringr", "tools", "lmer4", "combinat", "ggpubr", "report",
#                    "rcompanion", "sjPlot", "glmmTMB", "DHARMa", "tidyverse"))

library(lmerTest)
library(emmeans)
library(glue)
library(ggeffects)
library(ggplot2)
library(sjPlot)
library(report)
library(rcompanion)
library(sjPlot)

library(DHARMa)
library(glmmTMB)
library(tidyverse)
library(performance)
library(ordinal)
```

## Prepare directories

Set up directories - since we're doing this locally (and mounting the server), 
we're going to use a different base_dir

```{r}

base_dir<-'/Volumes/FinnLab/tommy/isc_asynchrony_behavior/derivatives/'
results_dir<-glue('{base_dir}results/behavioral/')
plots_dir <- glue('{base_dir}plots/final/')

```

## Functions

```{r}
# Function to convert specified columns to factors
convert_columns_to_factors <- function(df, columns_to_convert) {
  # Check if all specified columns exist in the dataframe
  if (!all(columns_to_convert %in% names(df))) {
    missing_cols <- columns_to_convert[!columns_to_convert %in% names(df)]
    stop(paste("The following columns do not exist in the dataframe:", 
               paste(missing_cols, collapse = ", ")))
  }
  
  # Convert each specified column to factor
  for (col in columns_to_convert) {
    df[[col]] <- factor(df[[col]])
  }
  
  return(df)
}

# Function to remove outliers based on IQR of a specific column and print the number of removed values
remove_outliers <- function(df, column) {
  if (!column %in% names(df)) {
    stop("Column not found in dataframe.")
  }
  if (!is.numeric(df[[column]])) {
    stop("Selected column is not numeric.")
  }

  initial_rows <- nrow(df)  # Store the initial number of rows

  # Calculate IQR bounds for the specified column
  Q1 <- quantile(df[[column]], 0.25)
  Q3 <- quantile(df[[column]], 0.75)
  IQR <- Q3 - Q1
  lower_bound <- Q1 - 1.5 * IQR
  upper_bound <- Q3 + 1.5 * IQR

  # Filter rows based on the specified column's bounds
  df_clean <- df[df[[column]] >= lower_bound & df[[column]] <= upper_bound, ]
  
  # Calculate and print number of rows removed
  removed_rows <- initial_rows - nrow(df_clean)
  cat("Number of rows removed due to outliers in column", column, ":", removed_rows, "\n")
  
  return(df_clean)
}


```

# Run human participant analysis

## Data loading

Load data and set factor columns 

```{r}

# Load the results file
results_fn<-glue('{results_dir}all-task_group-analyzed-behavior_human-lemmatized.csv')
df_results <- read.csv(results_fn)

# Set the variable types
factor_columns <- c('modality', 'ground_truth', 'word_index', 'top_pred', 'task')
df_results <- convert_columns_to_factors(df_results, factor_columns)

# subject model for accuracy --> test if audio > text
audio_greater_text <- c(1,-1)
contrasts(df_results$modality) <- cbind(audio_greater_text)
```

## Population-level Binary accuracy

### Logistic regression 

Classify accuracy based on modality. Conduct a contrast of audio > text

```{r}

binary_accuracy_formula <- 'accuracy ~ modality + (1|word_index:task)'

# Fit the logistic mixed-effects model
model <- glmer(binary_accuracy_formula, 
               data = df_results, 
               family = binomial(link = "logit"))

tab_model(model,show.se=TRUE, show.stat=TRUE, show.r2=TRUE)
summary (model)
report (model)
```

### Prosody interaction

Rerun the logistic regression with the mean prosody included as an interaction effect. 

```{r}

binary_accuracy_formula <- 'accuracy ~ modality * prominence_mean + (1|word_index:task)'

# Fit the logistic mixed-effects model
model <- glmer(binary_accuracy_formula, 
               data = df_results, 
               family = binomial(link = "logit"))

tab_model(model, show.se=TRUE, show.stat=TRUE, show.r2=TRUE)
summary (model)
```

### Plot interaction model

```{r}

plot_model(model, type="pred", terms=c("prominence_mean", "modality"))

```


## Continuous accuracy

### Linear mixed-effects model

Predict accuracy based on modality 

```{r}

continuous_accuracy_formula <- 'fasttext_top_word_accuracy ~ modality + (1|word_index:task)'

# Fit the logistic mixed-effects model
model <- lmer(continuous_accuracy_formula, data=df_results)

tab_model(model, show.se=TRUE, show.stat=TRUE, show.r2=TRUE)
summary (model)
report(model)
```

### Include prosody within model

```{r}

continuous_accuracy_formula <- 'fasttext_top_word_accuracy ~ modality * prominence_mean + (1|word_index:task)'

# Fit the logistic mixed-effects model
model <- lmer(continuous_accuracy_formula, data=df_results)

tab_model(model, show.se=TRUE, show.stat=TRUE, show.r2=TRUE, show.df=TRUE)
summary (model)

plot_model(model, type='pred', terms=c('prominence_mean', 'modality'))
```

# Supplementary analyses

## Leakage filtering stat check

```{r}

# First filter by pvalue less than 0.05 uncorrected --> test absence of effect
df_results_filtered <- subset(df_results, df_results$barnard_pvalue >= 0.05)

# subject model for accuracy --> test if audio > text
audio_greater_text <- c(1,-1)
contrasts(df_results_filtered$modality) <- cbind(audio_greater_text)

```

### Logistic regression 

Classify accuracy based on modality. Conduct a contrast of audio > text

```{r}

binary_accuracy_formula <- 'accuracy ~ modality + (1|word_index:task)'

# Fit the logistic mixed-effects model
model <- glmer(binary_accuracy_formula, 
               data = df_results_filtered, 
               family = binomial(link = "logit"))

tab_model(model,show.se=TRUE, show.stat=TRUE, show.r2=TRUE)
summary (model)
report (model)

```

### Prosody interaction

Rerun the logistic regression with the mean prosody included as an interaction effect. 

```{r}

binary_accuracy_formula <- 'accuracy ~ modality * prominence_mean + (1|word_index:task)'

# Fit the logistic mixed-effects model
model <- glmer(binary_accuracy_formula, 
               data = df_results_filtered, 
               family = binomial(link = "logit"))

tab_model(model, show.se=TRUE, show.stat=TRUE, show.r2=TRUE, show.df=TRUE)
summary (model)
```

## Continuous accuracy

### Linear mixed-effects model

Predict accuracy based on modality 

```{r}

continuous_accuracy_formula <- 'fasttext_top_word_accuracy ~ modality + (1|word_index:task)'

# Fit the logistic mixed-effects model
model <- lmer(continuous_accuracy_formula, data=df_results_filtered)

tab_model(model, show.se=TRUE, show.stat=TRUE, show.r2=TRUE, show.df=TRUE)
summary (model)
report(model)
```

### Include prosody within model

```{r}

continuous_accuracy_formula <- 'fasttext_top_word_accuracy ~ modality * prominence_mean + (1|word_index:task)'

# Fit the logistic mixed-effects model
model <- lmer(continuous_accuracy_formula, data=df_results_filtered)

tab_model(model, show.se=TRUE, show.stat=TRUE, show.r2=TRUE, show.df=TRUE)
summary (model)

plot_model(model, type='pred', terms=c('prominence_mean', 'modality'))
```

## Reaction times

### GLM with log transform 

```{r}

audio_greater_text <- c(1,-1)
contrasts(df_results$modality) <- cbind(audio_greater_text)

human_rt_formula <- 'average_rt ~ modality * prominence_mean + (1|task/word_index)'

df_results_no_outliers <- remove_outliers(df_results, 'average_rt')

# Fit the linear mixed-effects model
model <- glmer(human_rt_formula, data = df_results_no_outliers, 
               family=Gamma(link='log'),
               control = glmerControl(optimizer ="Nelder_Mead"))

tab_model (model)
summary (model)

# plot_model(model, type='pred', terms=c('prominence_mean', 'modality'), show.data = TRUE)

```

### Try beta distribution

```{r}

library(GLMMadaptive)

normalize <- function(x){(x-min(x))/(max(x)-min(x))}

df_results$fasttext_top_word_distance <- normalize(1 - df_results$fasttext_top_word_accuracy)
df_results[df_results$fasttext_top_word_distance == 1,]$fasttext_top_word_distance <- 0.999


# Fit zero-one inflated beta mixed model
model <- glmmTMB(
  fasttext_top_word_distance ~ modality  + (1|word_index),  # Main beta component
  ziformula = ~  modality + (1|word_index),  # Zero-inflation component
  data = df_results,
  family = beta_family(),
)

tab_model(model)

check_model(model)
```
